Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-02-15 automatica/paper.tex
Start merge of CAV 18 sections

%\newcommand{\tool}{{\sf DSSynth}\xspace}
\newcommand{\mat}[1]{\boldsymbol{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{proof}{Proof}

@@

@@

%-------------------------------
\subsection{Overapproximating continuous time dynamics using discrete models}
\label{sec:cont_aa}
%-------------------------------

Thus far we wave discussed the Abstract Accelertion of Discrete Time
systems.  However, our plant models derive from the discretization of
continuous time systems.  We therefore seek to establish a sound
overapproximation which does not only encompass the selected discretization,
but any chosen discretization in the time domain.

A dynamical system is a system in which a function describes the progression
of a state over time.  In a continuous domain with linear dynamics, it is
described by a first order Ordinary Differential Equation (ODE).
%
\begin{equation}
\dot{x}(t)=\mat{A}_o\vec{x}(t)+\mat{B}_o\vec{u}(t) +\vec{w}(t)
\label{eq:dynamical}
\end{equation}
%
where $\vec{w}$  is a continuous zero-mean white noise source with
covariance $N ( 0 , Q )$ Furthermore, a control system may have a derived
output that is a linear combination of its states and inputs, which may
restricts the observability of the statespace from the output space.
%
\begin{equation}
\vec{y}(t)=\mat{C}_o\vec{x}(t)+\mat{D}_o\vec{u}(t)
\end{equation}
%
Discretization of a continuous dynamical system turns the ODE into a
difference equation, assuming zero-order hold for the input $\vec{u}$ and
continuous integration for the noise $\vec{w}$, to
%
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k + \vec{w}_k\\
y_k &= \mat{C}_d \vec{x}_ k + \mat{D}_d \vec{u}_ k 
\end{align}
with covariance for $\vec{w}_k$, $N ( 0 , Q_d )$
where
\begin{align}
\label{eq:discretize}
\mat{A}_d &= e^{\mat{A}_o T_s} = \mathcal{L}^{-1} { ( s \mat{I} - \mat{A}_o )^{-1} }_{t = T_s}\\
\mat{B}_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} dt\ \mat{B}_o = \mat{A}_o^{-1} ( \mat{A}_d - \mat{I} ) \mat{B}_o\\
\mat{C}_d &= \mat{C}_o\\
\mat{D}_d &= \mat{D}_o \\
Q_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} Q e^{\mat{A}_o^T t} dt
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT)=x_k \wedge y(kT) = y_k, \forall k
\end{align*}

%-------------------------------
 \subsubsection{A discrete time model using reals}
 \label{sec:real_aa}
%-------------------------------

 \begin{theorem}
 The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0 : \dot{\vec{x}}=\mat{A}\vec{x} \wedge \mat{A}=\mat{S}\mat{J}\mat{S}^{-1}\nonumber\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \hdots  & s_i\frac{T^{p-i}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}\nonumber\\
 s_i&=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.
 \label{eq:continuous_tube_dyn}
 \end{align}
 where $\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x} : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
 \begin{proof}
 Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretized with sample time $T_1$ is ruled by the formula
 $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
\begin{align}
 e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
\end{align} 
\begin{align} 
 e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}(\mat{S}\mat{J}\mat{S}^{-1})^k
 =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}
 =\mat{S}e^{\mat{J}}\mat{S}^{-1}
\end{align} 
\begin{equation}
\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
 \end{equation}

 Let $\mat{J}$ be a diagonal matrix, such that 
 $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
 Let us now take a different sample rate $T_2$, such that 
 $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
 We can then say that 
 \begin{equation}
 {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
 \end{equation}
 This proves the theorem for $gm(\lambda_i)=1$.
 We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
 referenced to the original eigenvalues modified by the sampling rate.
  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix. Then
\begin{align}
 \mat{J}_1=&\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k\nonumber\\
 =&\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
 \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
& &  &\lambda_i^k \\
\end{array} \right ] T_1^k
\end{align}
%
The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%
\begin{align}
\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
\end{align}
which completes the proof for $gm(\lambda_i)>1$.
\end{proof}

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
 \begin{proof}
 We begin by expanding the equation
 \begin{align}
 \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\\
 \vec{x}_T'&=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u} = \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}\nonumber
 \end{align}
 where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
 \end{proof}
 \begin{theorem}
Let $\mat{A}_b$ be a spherical overapproximation of $\mat{A}$ and $U_b$ a spherical overapproximation of $U$ centered at $\vec{u}_c$
\begin{align}
\mat{A}_b&=\mat{S} \mat{J}_b \mat{S}^{-1}\\
\mat{J}_{bs}&=\lambda_{bs}\mat{I} \wedge \lambda_{bs}=|| \mat{J}_s ||=
\max_{\vec{u} \subseteq U_b'} \frac{|| \mat{J}_s\vec{u}_s ||}{||
  \vec{u}_s ||}\nonumber
\label{eq:induced}
\end{align}
where $|| \mat{J}_s ||$ is the induced matrix norm, $s$ indicates the index of each Jordan block, and $\vec{u}_s$ is the corresponding subvector of $\vec{u}$ (\emph{ie} the subspace of $\vec{u}$ on which $\mat{J}_s$ operates).
Directly from the definition of the matrix norm, we have 
\begin{equation}
||\mat{J}_s\vec{u}_s|| \leq ||\mat{J}_s||\,||\vec{u}_s||, 
\label{eq:overinduced}
\end{equation}
and $\mat{J}_{b} U_b$ is the smallest ball that over-approximates $\mat{J}  U_d$.

The expression
 \begin{align}
 \label{eq:contu_reachset}
X_T^\sharp &=\mat{A}^{-1}X_T'\supseteq X_T\\
X_T'&=\mat{A}\mat{A}_TX_0 \oplus (\mat{I}-\mat{A}_{bT})\mat{B}U_b \oplus (\mat{I}-\mat{A}_T)\mat{B}\vec{u}_c\nonumber
 \end{align}
 is an overapproximation of the reach set  at time $t=T$ of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t) + \mat{B}\vec{u}(t)$ given $\vec{x}(0) \in X_0$ and $\vec{u}(t) \in U$.
 \end{theorem}
\begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then 
 \begin{equation}
 \tiny
 X_k=A_d^kX_0 \oplus (\mat{I}-\mat{A}_{bd}^k)(\mat{I}-\mat{A}_{bd})^{-1}\mat{B}U_b \oplus (\mat{I}-\mat{A}_d^k)(\mat{I}-\mat{A}_d)^{-1}\mat{B}\vec{u}_c\nonumber\\
 k \in (0,\infty)
 \end{equation}
  is a witness of the system at time $kT_d$, where $\mat{A}_{bd}$ is a spherical overapproximation of $\mat{A}_d$
\end{corollary}
 \begin{proof}
Let $U'=\mat{B}U$, with $\vec{u}'$ its geometrical centre and $U_b'$ the smallest sphere centered at $\vec{u}'$ containing all points in $U'$.
Then $U'\subseteq U_b' \oplus \vec{u}_c' \Rightarrow (\mat{I}-\mat{A}_T)U' \subseteq (\mat{I}-\mat{A}_T)U_b' \oplus (\mat{I}-\mat{A}_T)\vec{u}_c'$. Since $ (\mat{I}-\mat{A}_{bT})U_b'\supseteq  (\mat{I}-\mat{A}_T)U_b'$, then $(\mat{I}-\mat{A}_T)U' \subseteq (\mat{I}-\mat{A}_{bT})U_b' \oplus (\mat{I}-\mat{A}_T)\vec{u}_c'$ thus proving the theorem for $\vec{u}(t)=\vec{u}_0 : t \in [0\ T]$.

For the continuous time interval, let us look at the solution for the differential equation at time $T$ with a non-deterministic input $\vec{u}$.
\begin{equation*}
\vec{x}_T=e^AT\vec{x}_0+\int_{t = 0}^{T_s} e^{\mat{A} t} \mat{B} \vec{u}(t) dt
\end{equation*}
We will first divide this equation into 
\begin{align*}
\vec{x}_T&=e^AT\vec{x}_0+\int_{t = 0}^{T} e^{\mat{A} t} \mat{B} \vec{u}_c dt +\int_{t = 0}^{T_s} e^{\mat{A} t} \mat{B} (\vec{u}(t)-\vec{u}_c) dt\\
&=A_T\vec{x}_0+B_T\vec{u}_c+\int_{t = 0}^{T} \mat{S}e^{\mat{J} t} \mat{S}^{-1}\mat{B} (\vec{u}(t)-\vec{u}_c) dt\\
&=A_T\vec{x}_0+B_T\vec{u}_c+\mat{S}\int_{t = 0}^{T} e^{\mat{J} t} \vec{u}_{j}'(t) dt\\
&=A_T\vec{x}_0+B_T\vec{u}_c+\mat{S}\int_{t = 0}^{T} \sum_{k=0}^\infty \frac{1}{k!} \mat{J}^k t^k \vec{u}_{j}'(t) dt\\
&=A_T\vec{x}_0+B_T\vec{u}_c+\mat{S}\sum_{k=0}^\infty \frac{1}{k!} \int_{t = 0}^{T} t^k \mat{J}^k \vec{u}_{j}'(t) dt
\end{align*}
It follows from equation \eqref{eq:overinduced} that $\mat{J}^k \vec{u}_{j}'(t) \subseteq \mat{J}_b^k U_{bj}'$ (where $U_{bj}'=\mat{S}^{-1}U_b'$), hence
\begin{align*}
&X_T \subseteq A_TX_0+B_T\vec{u}_c+\mat{S}\sum_{k=0}^\infty \frac{1}{k!} \int_{t = 0}^{T} t^k \mat{J}_b^k U_{bj}' dt\\
\Rightarrow &X_T \subseteq A_TX_0+B_T\vec{u}_c+\mat{S} \int_{t = 0}^{T} \sum_{k=0}^\infty \frac{1}{k!}\mat{J}_b^k t^k U_{bj}' dt\\
\Rightarrow &X_T \subseteq A_TX_0+B_T\vec{u}_c+\int_{t = 0}^{T_s} e^{\mat{A}_b t} U_b' dt\\
\Rightarrow &X_T \subseteq A_TX_0+B_T\vec{u}_c+ \mat{A}^{-1}(\mat{I}-\mat{A}_{bT}) \mat{B} U_b
\end{align*}
Multiplying by $\mat{A}$ we get $X_T'$ as in \eqref{eq:contu_reachset}, thus proving the theorem.
\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The final case we analyse here is that of hybrid time systems, where the plant dynamics evolve in continuous time
while the feedback dynamics evolve in discrete time.

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
 \end{align}
 with $\mat{A}_t, \mat{B}_t$ as per \eqref{eq:continuous_tube_dyn} and $\mat{A}_d, \mat{B}_d$ as per equation \eqref{eq:discretize} is a witness of the system
 \begin{align}
 \dot{\vec{x}}_{t} &= \mat{A}\vec{x}+\mat{B}\vec{u}_t\nonumber\\
 \vec{u}_t=\mat{K}\vec{x}_k  : kT_s \leq t < (k+1)T_s
 \end{align}
 \end{theorem}

\begin{proof}
Let $\mat{K}$ be a feedback controller such that $\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$. Let the output of
the feedback system be a zero order holder such that $\vec{u}(t+kT_s)=\vec{u}(kT_s) : 0 \leq t < T_s$.
The closed loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
\begin{equation}
\vec{x}_{k,t}=\mat{A}_t\vec{x}_k-\mat{B}_t\mat{K}\vec{x}_k = (\mat{A}_t-\mat{B}_t\mat{K})\vec{x}_k : 0 \leq t < T_s
\end{equation}
We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
\begin{align}
\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
\end{align}
which accelerating results becomes
\begin{align}
\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
\vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
\end{align}
\end{proof}

>>>>

%\newcommand{\tool}{{\sf DSSynth}\xspace}
\newcommand{\mat}[1]{\boldsymbol{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{proof}{Proof}

@@

%-------------------------------
\subsection{State-space representations of LTI models}
\label{sec:model}
%-------------------------------

%% A dynamical system is a system in which a function describes the progression
%% of a state over time.  In a continuous domain with linear dynamics, it is
%% described by a first order Ordinary Differential Equation (ODE).
%% \begin{equation}
%% \dot{x}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)% +\vec{w}(t)
%% \label{eq:dynamical}
%% \end{equation}

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable~\cite{Astrom08,
astrom1997computer}:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad 
\vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q,\,
\mat{A} \in \mathbb{R}^{p \times p},\,
\mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

%%-------------------------------
%\subsection{Overapproximating continuous time dynamics using discrete models}
%\label{sec:cont_aa}
%%-------------------------------
%
%Thus far we wave discussed the Abstract Accelertion of Discrete Time
%systems.  However, our plant models derive from the discretization of
%continuous time systems.  We therefore seek to establish a sound
%overapproximation which does not only encompass the selected discretization,
%but any chosen discretization in the time domain.
%
%A dynamical system is a system in which a function describes the progression
%of a state over time.  In a continuous domain with linear dynamics, it is
%described by a first order Ordinary Differential Equation (ODE).
%%
%\begin{equation}
%\dot{x}(t)=\mat{A}_o\vec{x}(t)+\mat{B}_o\vec{u}(t)
%\label{eq:dynamical}
%\end{equation}
%%
%Furthermore, a control system may have a derived
%output that is a linear combination of its states and inputs, which may
%restricts the observability of the statespace from the output space.
%%
%\begin{equation}
%\vec{y}(t)=\mat{C}_o\vec{x}(t)+\mat{D}_o\vec{u}(t)
%\end{equation}
%%
%Discretization of a continuous dynamical system turns the ODE into a
%difference equation, assuming zero-order hold for the input $\vec{u}$, to
%%
%\begin{align}
%\label{eq:discretization}
%\vec{x}_{k+1} &= \mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
%y_k &= \mat{C}_d \vec{x}_ k + \mat{D}_d \vec{u}_ k 
%\end{align}
%where
%\begin{align}
%\label{eq:discretize}
%\mat{A}_d &= e^{\mat{A}_o T_s} = \mathcal{L}^{-1} { ( s \mat{I} - \mat{A}_o )^{-1} }_{t = T_s}\\
%\mat{B}_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} dt\ \mat{B}_o = \mat{A}_o^{-1} ( \mat{A}_d - \mat{I} ) \mat{B}_o\\
%\mat{C}_d &= \mat{C}_o\\
%\mat{D}_d &= \mat{D}_o \\
%\end{align}
%and $T_s$ is the sample time. Then
%\begin{align*}
%x(kT)=x_k \wedge y(kT) = y_k, \forall k
%\end{align*}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$, we must be able to ensure that
certain properties of the plant hold at any given point in time.  In order
to do so, we accelerate the dynamics of the plant according to the model
semantics in Sec.~\ref{sec:model} and compute the state
$\vec{x}_T$ at a given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

%\note{ [add?] In this section we work with fully observed models, namely where $y = x$. } 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
Given a specific time $T$, the state $\vec{x}(t=T)$ can be expressed as a
linear function of $\vec{x}_0$ as shown below.  We assume that the system
has no inputs, so its dynamics are described by
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$.

\begin{lemma}
The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, and $\mat{S}$ are the generalized eigenvectors of $\mat{A}$,
evaluated at time $T$ is
%
\begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn2}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
\end{align}
%
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.
%
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T. 
If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0 \text{ with } T \in \mathbb{R}^+_0$ is the state of the system at time $t=T$. 
\end{lemma}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

Next, we are interested in studying models with parametric inputs, which are
described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where
$\forall t\geq 0.\,\vec{u}(t)=\vec{u}$.  For such models, a state $\vec{x}_T$
is computed next.
%
\begin{lemma}
%
The state of the model
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a
parametric input $\vec{u}$ is
%
\begin{align}
\vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
\label{eq:continuous_tube_param}
\end{align}
%
where $\forall t \leq T.\, \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in
\eqref{eq:continuous_tube_dyn2} \footnote{When $\mat{A}$ is not invertible,
we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\
0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$.}. 
If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that
$A_d=e^{\mat{A} T_d}$, then
$\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u})
\text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} =
e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the
system at time $T$.
%
\end{lemma}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Accelerated dynamics for models with a sampled-time feedback}\label{sec:real_discrete_feedback_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The final case we analyze is that of hybrid-time models, where the plant
dynamics evolve in continuous time while the feedback dynamics evolve in
discrete time.  A~new input value is produced only at every sample time but
implemented over the continuous-time plant dynamics.
%
\begin{theorem}
The expression
%
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
%
with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in
\eqref{eq:continuous_tube_param}, and where $\mat{A}_d=\mat{A}_{T_s}$,
$\mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with
sample time $T_s$, is the state of the model
%
\begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s \,. 
\end{align*}
%
\end{theorem}
\begin{proof}
%
Let $\mat{K}$ be a feedback controller such that
$\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$.  Let the output of the
feedback system be a zero order holder such that
$\vec{u}(t'+kT_s)=\vec{u}(kT_s), \text{ for } 0 \leq t' < T_s$.  The closed
loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%
\begin{align*}
\vec{x}_{k,t'}=\mat{A}_{t'}\vec{x}_k-\mat{B}_{t'}\mat{K}\vec{x}_k = (\mat{A}_{t'}-\mat{B}_{t'}\mat{K})\vec{x}_k,
\end{align*}
where $0 \leq t' < T_s$
We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
\begin{align*}
\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
\end{align*}
which accelerating results and setting $T=kT_s+t'$ becomes
\begin{align}
\label{eq:feedback_sampled_cont}
\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
\vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
\label{eq:feedback_cont}
\end{align}
\end{proof}
